---
title: "p8105_hw2_np2619"
author: "Nidhi Patel"
date: "9/28/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1
###### Clean up Mr. Trash Wheel
```{r Prob_1_trash}
trash = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = "A2:N406") %>%
janitor::clean_names() %>%
drop_na(dumpster) %>%
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
  )


```

###### Clean up precipitation 2017 and 2018

Import datasets + clean a bit

```{r Precipitation}
precip17 = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
        range = "A2:B14") %>%
janitor::clean_names() %>% 
mutate(year = "2017") %>% 
  mutate(month = month.name)

precip18 = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    range = "A2:B14") %>%
janitor::clean_names() %>% 
mutate(year = "2018") %>% 
  mutate(month = month.name)

```

###### Join datasets

```{r}
precipitation = 
  bind_rows(precip17, precip18)
```

The Mr. Trash Wheel data has `r count(trash)` observations after cleaned.  The median number of sports balls in a dumpster in 2017 is `r median(pull(trash, sports_balls))`. 

The precipitation data set looks at amount of rainfall in inches per month for 2017 and 2018.  In 2017, the average precipitation is `r mean(pull(precip17, total))` with the min of `r min(pull(precip17, total))` and max of `r max(pull(precip17, total))`. In 2018, the average monthly precipitation is `r mean(pull(precip18, total))` with a min of `r min(pull(precip18, total))` and max of `r max(pull(precip18, total))`.  For the two years the average precipitation is `r mean(pull(precipitation, total))` with a total of `r count(precipitation)` observations.  The total precipitation in 2018 is `r sum(pull(precip18, total))`.


## Problem 2
###### Tidy Transit data
```{r}
transit = 
  read_csv("./NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line:entry, vending, ada) %>% 
  mutate(entry = as.logical(recode(entry, "YES" = "TRUE", "NO" = "FALSE"))) %>% 
  mutate_at(vars(route8:route11), as.character) %>% 
  pivot_longer(
    route1:route11,
    names_to = "route",
    names_prefix = "route",
    values_to = "rt_name"
  ) %>% 
  relocate(route, rt_name) %>% 
  drop_na(rt_name)

```

First, to tidy this data, I cleaned names and converted the character 'entry' vector to logical.  This dataset had many routes that can be combined into one column, but first all the routes have to match and route8-route11 were converted to characters to match route1-route7. I used `pivot_longer` to tidy the data for easier analysis.  In doing so, multiple route columns were combined to one, with their values as 'rt_name'.  After this step, I dropped the NA values of rt_names to remove excess rows from creating a singular 'route' variable.  In total, there are 4,270 observations and 10 vectors (4,270 x 10).  Variables in the dataset include route, rt_name (the subway line at the route), line, station name, station latitude, station longitude, entrance type, entry, vending, and ADA compliance.  

###### Questions

```{r}
distinct(transit, rt_name, line) %>% 
  count()

count(transit, ada == "TRUE")

filter(transit, vending == "NO") %>% 
  count()

filter(transit, vending == "NO", entry == "TRUE") %>% 
  count()

prop = 139/447

```


There are 259 distinct stations and 1616 stations are ADA compliant.  The proportion of stations without vending that allow entrance can be found by the number of stations without vending that allow entrances/exits (139) over the number of stations without vending(447). This is `r prop * 100`%



## Problem 3

Merging data using Year and month as keys across the 3 datasets. 

```{r}
pols = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
mutate(
  year = as.integer(year),
  month = as.integer(month),
  day = as.integer(day)
  )
 
month_df =
  tibble(
    month = 1:12,
    month_name = month.name
  )
 
polstidy = 
  left_join(pols, month_df, by = "month") %>% 
  relocate(year, month_name, prez_gop, prez_dem) %>% 
  pivot_longer(
    prez_gop:prez_dem, 
    names_to = "president",
    names_prefix = "prez_",
    values_to = "partyaffil") %>% 
  select(year:month_name, gov_gop:partyaffil)
  
  
   
  

```

