---
title: "p8105_hw2_np2619"
author: "Nidhi Patel"
date: "9/28/2020"
output: github_document
---

```{r setup}
library(tidyverse)
library(readxl)
```


## Problem 1
###### Clean up Mr. Trash Wheel
```{r Prob_1_trash}
trash = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "Mr. Trash Wheel",
    range = "A2:N406") %>%
janitor::clean_names() %>%
drop_na(dumpster) %>%
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)
  )


```

###### Clean up precipitation 2017 and 2018

Import datasets + clean a bit

```{r Precipitation}
precip17 = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2017 Precipitation",
        range = "A2:B14") %>%
janitor::clean_names() %>% 
mutate(year = "2017") %>% 
  mutate(month = month.name)

precip18 = 
  read_excel(
    "./Trash-Wheel-Collection-Totals-8-6-19.xlsx", 
    sheet = "2018 Precipitation",
    range = "A2:B14") %>%
janitor::clean_names() %>% 
mutate(year = "2018") %>% 
  mutate(month = month.name)

```

###### Join datasets

```{r}
precipitation = 
  bind_rows(precip17, precip18)
```

The Mr. Trash Wheel data has `r count(trash)` observations after cleaned.  The median number of sports balls in a dumpster in 2017 is `r median(pull(trash, sports_balls))`. 

The precipitation data set looks at amount of rainfall in inches per month for 2017 and 2018.  In 2017, the average precipitation is `r mean(pull(precip17, total))` with the min of `r min(pull(precip17, total))` and max of `r max(pull(precip17, total))`. In 2018, the average monthly precipitation is `r mean(pull(precip18, total))` with a min of `r min(pull(precip18, total))` and max of `r max(pull(precip18, total))`.  For the two years the average precipitation is `r mean(pull(precipitation, total))` with a total of `r count(precipitation)` observations.  The total precipitation in 2018 is `r sum(pull(precip18, total))`.


## Problem 2


